### 2025-05-17 teaching.qmd

---
title: "Teaching"
---

# Teaching

## 2020 Fall

- Statistical Models: Theory and Application  [[STAT 215A](https://classes.berkeley.edu/content/2019-Fall-STAT-215A-001-LEC-001)]

## 2019 Fall

- Statistical Models: Theory and Application  [[STAT 215A](https://classes.berkeley.edu/content/2019-Fall-STAT-215A-001-LEC-001)]

## 2019 Spring

- [Modern Statistical Prediction and Machine Learning](https://stat154.github.io/logistics/)  [[STAT 154](https://classes.berkeley.edu/content/2019-Spring-STAT-154-001-LEC-001)]