### 2025-07-20 index.qmd
Update website content and layout
Changed the page footer in _quarto.yml to center the copyright notice for 2025. Revised the 'About' page in about.qmd to enhance clarity and organization, including a new section for research themes and updated recent publications. Improved the layout of index.qmd to utilize a two-column format for better presentation of Professor Bin Yu's information and contact details. Enhanced styles.css to support the new column layout and added styling for year entries in publication lists.

---
title: Home
title-block: false
format:
  html:
    css: styles.css
---

:::: {.columns}

::: {.column width="40%"}
![](images/binyu.jpg)
:::

::: {.column width="60%"}
## Professor Bin Yu

Chancellor's Distinguished Professor of Statistics  
Statistics Department Chair, 2009 - 2012  
Professor of Electrical Engineering and Computer Sciences, by courtesy

## Contact

Address: 367 Evans Hall #3860, Berkeley, CA 94720  
Phone: 510-642-2781  
Fax: 510-642-7892  
Email: [binyu@berkeley.edu](mailto:binyu@berkeley.edu)
:::

::::

## Recent Publications, Talks & News

- <span class="year-entry">2025</span>[PCS workflow for veridical data science in the age of AI](https://www.stat.berkeley.edu/~yugroup/pubs/pcs_workflow-6.pdf) - Framework for responsible AI development and deployment
- <span class="year-entry">2025</span>[Veridical Data Science in Biology workshop](https://www.eventbrite.com/e/veridical-data-science-for-biology-2025-tickets-1384456339179), UC Berkeley, July 11 - Applying trustworthy data science to biological research
- <span class="year-entry">2025</span>[Rome Workshop on Veridical Data Science](https://www.integreat.no/events/public-events/workshops/veridical-data-science.html), June 20 - International workshop on responsible data analysis
- <span class="year-entry">2024</span>[Veridical data science and medical foundation models](https://arxiv.org/abs/2409.10580) - Guidelines for trustworthy medical AI systems
- <span class="year-entry">2024</span>[Mechanistic Interpretation through Contextual Decomposition in Transformers](https://arxiv.org/pdf/2407.00886) - Method for understanding transformer decision-making processes
- <span class="year-entry">2024</span>[LoRA+: Efficient Low Rank Adaptation of Large Models](https://arxiv.org/abs/2402.12354) - Improved technique for fine-tuning large language models
- <span class="year-entry">2024</span>[VDS book review](https://hdsr.mitpress.mit.edu/pub/5322mn1e/release/3) in Harvard Data Science Review - Praise for pedagogical approach to responsible data science
- <span class="year-entry">2024</span>[Berkeley-Stanford Workshop on Veridical Data Science](https://na.eventscloud.com/website/69057/) ([videos](https://www.youtube.com/playlist?list=PLKW2Azk23ZtSrg4jwLSioaV1bwmFnQ1Bx)) - Inaugural academic collaboration on trustworthy AI
- <span class="year-entry">2024</span>[Veridical Data Science book](https://mitpress.mit.edu/9780262049191/veridical-data-science/) released ([online](http://vdsbook.com/)) - Comprehensive guide to responsible data analysis and decision making
- <span class="year-entry">2023</span>[Explaining black box text modules in natural language with language models](https://arxiv.org/abs/2305.09863) - Natural language explanations for neural network text processing
- <span class="year-entry">2023</span>[Diagnosing transformers: illuminating feature space for clinical decision-making](https://arxiv.org/abs/2305.17588) - Interpretability methods for medical transformer applications
- <span class="year-entry">2023</span>[COPSS Distinguished Award and Lecture](https://ww2.amstat.org/meetings/jsm/2023/webcasts/index.cfm), JSM Toronto - Recognition for contributions to statistical science and AI

[View all papers →](papers.qmd)