### 2025-07-20 people.qmd
Update 'People' section in people.qmd to relocate Yeshwanth Cherapanamjeri's entry
Moved Yeshwanth Cherapanamjeri's listing from the current post-docs section to the former post-docs section for better organization and clarity.

---
title: "People"
---

# People

## Current Members (Role, Program)

- Abhineet Agarwal (Ph.D. Student, Statistics)
- Soufiane Hayou (Post-doc)
- Jakob Heiss (Post-doc)
- Aliyah Hsu (Ph.D. Student, Statistics)
- Yaxuan Huang (Ph.D. Student, Statistics)
- Nathan Benjamin McNaughton (Undergraduate Student)
- Robin Netzorg (Ph.D. Student, Statistics)
- Anthony Ozerov (Ph.D. Student, Statistics)
- Zachary Rewolinski (Ph.D. Student, Statistics)
- Omer Ronen (Ph.D. Student, Statistics)
- Jingfeng Wu (Post-doc)
- Chengzhong Ye (Ph.D. Student, Statistics)
- Zeyu Yun (Ph.D. Student, Statistics)
- Austin Zane (Ph.D. Student, Statistics)

## Former Ph.D. Students (Graduation Year)

- Corine Elliot (2024) - Statistical Scientist, Berry Consultants
- James Duncan (2024) - Climate ML Engineering, Allen Institute for AI (AI2)
- Tiffany Tang (2023) - Clare Boothe Luce Assistant Professor, University of Notre Dame
- Dennis Shen (2023) - Assistant Professor, USC
- Spencer Frei (2023) - Assistant Professor, UC Davis
- Rebecca L. Barter (2022) - Research Assistant Professor, University of Utah
- Luiz Chamon (2022) - Research Group Leader, University of Stuttgart
- Angela Zhou (2022) - Assistant Professor, University of Southern California
- Briton Park (2022) - Citadel Securities
- Chandan Singh (2022) - Microsoft Research
- Keyan Nasseri (2022) - Google
- Wooseok Ha (2022) - AWS
- Yan Shuo Tan (2022) - Assistant Profressor, Department of Statistics and Data Science, National University of Singapore
- Xiao Li (2021) - Voleon
- Raaz Dwivedi (2021) - Postdoctoral scholar, MIT / Harvard
- Yu Wang (2020) - Two Sigma
- Merle Behr (2020) - Scientific Expert, Bayer AG
- Nicholas Altieri (2020) - Genentech
- Yuansi Chen (2019) - Assistant Professor, Department of Statistical Sciences, Duke University
- Karl Kumbier (2019) - Postdoctoral Researcher, UC San Francisco
- Jamie Murdoch (2019) - Co-founder / President, clientelligent.ai
- Soren Kunzel (2019) - Data Scientist, Citadel
- Simon Walter (2019) - Data Scientist, Bridgewater
- Sujayam Saha (2018) - Data Scientist, Google
- Reza Abbasi-Asl (2018) - Scientist, Allen Institute for Brain Science
- Christine Kuang (2017) - Data Scientist, Facebook
- Hanzhong Liu (2016) - Assistant Professor, Department of Statistics, Tsinghua University
- Sumanta Basu (2016) - Assistant Professor, Department of Statistics, Cornell University
- Sivaraman Balakrishnan (2016) - Assistant Professor, Department of Statistics, Carnegie Mellon University
- Adam Bloniarz (2016) - Data Scientist, Google
- Siqi Wu (2016) - Data Scientist, Citadel
- Hongwei Li (2015) - Senior Data Scientist, UBER
- Antony Joseph (2014) - Staff Data Scientist, Walmart eCommerce
- Taesup Moon (2013) - Research Staff Member, Samsung Advanced Institute of Technology
- Yuval Benjamini (2013) - Stein fellow, Statistics department, Stanford University
- Julien Mairal (2012) - Research Scientist, INRIA
- Yueqing Wang (2012) - Research Scientist, Google
- Hau-tieng Wu (2012) - Assistant Professor, Mathematics Department, University of Toronto
- Garvesh Raskutti (2012) - Assistant Professor, Department of Statistics, Univ. Wisconsin-Madison
- Luke W. Miratrix (2012) - Assistant Professor, Department of Statistics, Harvard University
- Harry Kim (2011) - Product Marketing Manager, Google
- Chinghway Lim (2011) - Assistant Professor, Department of Statistics and Applied Probability, National University of Singapore
- Karl Rohe (2011) - Assistant Professor, Department, University of Wisconsin-Madison.
- Kyle Jia (2010) - Assistant Professor, Peking University, China.
- David Purdy (2009) - Data Scientist, Uber
- Pradeep Ravikumar (2009) - Associate Professor, Department of Computer Sciences, University of Texas, Austin.
- Vince Vu (2009) - Assistant Professor, Department of Statistics, The Ohio State University
- Ethan Anderes (2008) - Assistant Professor, Department of Statistics, University of California, Davis
- Guilherme Rocha (2008) - Assistant Professor, Department of Statistics, Indiana University, Bloomington
- Peng Zhao (2006) - CEO of Citadel Securities, Citadel Investment Group in Chicago, IL.
- Nicolai Meinshausen (2006) - Professor, Department of Statistics, ETH Zurich
- Tao Shi (2005) - Quantitative Researcher, Citadel Investment Group in Chicago, IL.
- Gang Liang (2004) - Quantitative Operations Associate, Bank of America
- Rebecka Jörnsten (2000) - Associate Professor, Mathematical Statistics, Chalmers University of Technology
- Grace Chang (2000) - Senior Manager of Product Development, Marketing Dept, MTN Irancell

## Former Post-docs, Visitors, and Associates (Starting Year)

- Yeshwanth Cherapanamjeri (2023)
- Arman Sabbaghi (2022)
- Yanjun Qi (2020)
- Ezequiel Smucler (2020)
- Laura Reiger (2019)
- Yanyan Lan (2018)
- Yahong Han (2016)
- Thibault Vatter
- Guoqiang Cai
- Trine Abrahamsen
- Xiaoling Lu
- Xiusheng Lu (2014)
- Xiangyu Chang (2012)
- Ling Hong (2012)
- Toshiyasu Matsushima (2012)
- Yanfeng Gu (2012) - Professor of Electronics Engineering, Harbin Institute of Technology
- Hai Zhang (2012)
- Rodolphe Jenatton (2010)
- Kenichi Hayashi (2010)
- Yangbo He (2010)
- Fei Wu (2010)
- Kei Kobayashi (2008)
- Xing Wang (2007)