### 2025-11-18 books/index.qmd
fixed book image link, added welcome info, and added 214 to teaching page

---
title: Books
---

# Books

## Veridical Data Science

<div style="display: flex; align-items: flex-start; margin-bottom: 2em;">

<img src="../images/veridical-data-science-cover.jpg" alt="Veridical Data Science Book Cover" style="float:left; margin:0 2em 1em 0; width:200px; border-radius:8px;" />

<div>

*Published by MIT Press, 2024*

*Veridical Data Science* presents the PCS (Predictability, Computability, and Stability) framework that prioritizes reproducibility, interpretability, and reliability in data analysis. 
The book covers essential topics including statistical modeling, machine learning, causal inference, and ethical considerations in data science. Through real-world case studies and practical examples, readers learn how to apply the PCS principles to ensure their analyses are not only accurate but also robust and interpretable.

- **[Web](https://vdsbook.com/)** (free online access)
- **[Printed](https://mitpress.mit.edu/9780262049191/veridical-data-science/)**

</div>

</div>