### 2025-11-18 index.qmd

---
title: Home
title-block: false
format:
  html:
    css: styles.css
---

:::: {.columns}

::: {.column width="40%"}
![](images/binyu.jpg)
:::

::: {.column width="60%"}
## Professor Bin Yu

CDSS Chancellor's Distinguished Professor of Statistics  
Electrical Engineering and Computer Sciences, and Center for Computational Biology  
Statistics Department Chair, 2009 - 2012

## Contact

Address: 367 Evans Hall #3860, Berkeley, CA 94720  
Phone: 510-642-2781  
Fax: 510-642-7892  
Email: [binyu@berkeley.edu](mailto:binyu@berkeley.edu)

:::

::::

## Welcome

Professor Bin Yu is the head of the Yu Group at Berkeley, which consists of
12-15 students and postdocs from Statistics and EECS. She was formally trained
as a statistician, but her research interests and achievements extend beyond
the realm of statistics. Together with her group,  Bin Yu has leveraged new
computational developments to solve important scientific problems by combining
novel and often interpretable statistical machine learning approaches with the
domain expertise of my many collaborators in neuroscience, genomics and
precision medicine. She also develops relevant theory to understand random
forests and deep learning for insight into and guidance for practice.  Her work
has been recongized by many awards. In particular, she was inducted to the
National Academy of Sciences in 2014 and to the American Academy of Arts and
Sciences in 2013. She delivered the IMS Wald Lectures and COPSS Distinguished
Achievement and Award Lecture (DAAL, formerly Fisher Award and Lecture) in 2023.

She and her team have developed the PCS framework for veridical data science (or
responsible, reliable, and transparent data analysis and decision-making). PCS
stands for predictability, computability and stability, and it unifies,
streamlines, and expands on ideas and best practices of machine learning and
statistics.

In order to augment empirical evidence for decision-making, they are
investigating statistical machine learning methods/algorithms (and associated
statistical inference problems) such as dictionary learning, non-negative matrix
factorization (NMF), EM and deep learning (CNNs and LSTMs), and heterogeneous
effect estimation in randomized experiments (X-learner). Their recent algorithms
include staNMF for unsupervised learning, iterative Random Forests (iRF) and
signed iRF (s-iRF) for discovering predictive and stable high-order interactions
in supervised learning, next generation tree-based methods (e.g. fast and
interpretable greedy-tree sums (FIGS) and hierarchical shrinked (HS) trees, and
RF+), contextual decomposition (CD), aggregated contextual decomposition (ACD),
and adaptive wavelet distillation (AWD) for interpretation of Deep Neural
Networks (DNNs).

## Recent Publications, Talks & News

- <span class="year-entry">2025</span>[PCS workflow for veridical data science in the age of AI](https://www.stat.berkeley.edu/~yugroup/pubs/pcs_workflow-6.pdf) - Framework for responsible AI development and deployment
- <span class="year-entry">2025</span>[Veridical Data Science in Biology workshop](https://www.eventbrite.com/e/veridical-data-science-for-biology-2025-tickets-1384456339179), UC Berkeley, July 11 - Applying trustworthy data science to biological research
- <span class="year-entry">2025</span>[Rome Workshop on Veridical Data Science](https://www.integreat.no/events/public-events/workshops/veridical-data-science.html), June 20 - International workshop on responsible data analysis
- <span class="year-entry">2024</span>[Veridical data science and medical foundation models](https://arxiv.org/abs/2409.10580) - Guidelines for trustworthy medical AI systems
- <span class="year-entry">2024</span>[Mechanistic Interpretation through Contextual Decomposition in Transformers](https://arxiv.org/pdf/2407.00886) - Method for understanding transformer decision-making processes
- <span class="year-entry">2024</span>[LoRA+: Efficient Low Rank Adaptation of Large Models](https://arxiv.org/abs/2402.12354) - Improved technique for fine-tuning large language models
- <span class="year-entry">2024</span>[VDS book review](https://hdsr.mitpress.mit.edu/pub/5322mn1e/release/3) in Harvard Data Science Review - Praise for pedagogical approach to responsible data science
- <span class="year-entry">2024</span>[Berkeley-Stanford Workshop on Veridical Data Science](https://na.eventscloud.com/website/69057/) ([videos](https://www.youtube.com/playlist?list=PLKW2Azk23ZtSrg4jwLSioaV1bwmFnQ1Bx)) - Inaugural academic collaboration on trustworthy AI
- <span class="year-entry">2024</span>[Veridical Data Science book](https://mitpress.mit.edu/9780262049191/veridical-data-science/) released ([online](http://vdsbook.com/)) - Comprehensive guide to responsible data analysis and decision making
- <span class="year-entry">2023</span>[Explaining black box text modules in natural language with language models](https://arxiv.org/abs/2305.09863) - Natural language explanations for neural network text processing
- <span class="year-entry">2023</span>[Diagnosing transformers: illuminating feature space for clinical decision-making](https://arxiv.org/abs/2305.17588) - Interpretability methods for medical transformer applications
- <span class="year-entry">2023</span>[COPSS Distinguished Award and Lecture](https://ww2.amstat.org/meetings/jsm/2023/webcasts/index.cfm), JSM Toronto - Recognition for contributions to statistical science and AI

[View all papers →](papers.qmd)