### 2025-11-18 index.qmd
edit to welcome and specified google scholar for papers after 2025

---
title: Home
title-block: false
format:
  html:
    css: styles.css
---

:::: {.columns}

::: {.column width="40%"}
![](images/binyu.jpg)
:::

::: {.column width="60%"}
## Professor Bin Yu

CDSS Chancellor's Distinguished Professor of Statistics  
Electrical Engineering and Computer Sciences, and Center for Computational Biology  
Statistics Department Chair, 2009 - 2012

## Contact

Address: 367 Evans Hall #3860, Berkeley, CA 94720  
Phone: 510-642-2781  
Fax: 510-642-7892  
Email: [binyu@berkeley.edu](mailto:binyu@berkeley.edu)

:::

::::

## Welcome

Professor Bin Yu leads the Yu Group at UC Berkeley, an interdisciplinary team in
Statistics and EECS dedicated to advancing machine learning, artificial
intelligence, and veridical data science. Her group develops efficient and
interpretable ML/AI methods and theory—ranging from iterative random forests
(iRF) and tree-based FIGS to LoRA+ for fine-tuning deep learning and CD-T and
SPEX for interpreting deep models. The Yu Group collaborates closely with domain
experts in medical AI, genomics, and neuroscience.

A member of both the National Academy of Sciences and the American Academy of
Arts and Sciences, she has delivered major keynotes, including the 2019 Breiman
Lecture at NeurIPS, 2023 IMS Wald Lectures and the COPSS DAAL Lecture (formerly
Fisher Lecture). She and her team pioneered the PCS framework (predictability,
computability, and stability) for veridical (truthful) data science (VDS),
which has become an influential guide for transparent, trustworthy data science
and AI.

## Recent Publications, Talks & News

- <span class="year-entry">2025</span>[PCS workflow for veridical data science in the age of AI](https://www.stat.berkeley.edu/~yugroup/pubs/pcs_workflow-6.pdf) - Framework for responsible AI development and deployment
- <span class="year-entry">2025</span>[Veridical Data Science in Biology workshop](https://www.eventbrite.com/e/veridical-data-science-for-biology-2025-tickets-1384456339179), UC Berkeley, July 11 - Applying trustworthy data science to biological research
- <span class="year-entry">2025</span>[Rome Workshop on Veridical Data Science](https://www.integreat.no/events/public-events/workshops/veridical-data-science.html), June 20 - International workshop on responsible data analysis
- <span class="year-entry">2024</span>[Veridical data science and medical foundation models](https://arxiv.org/abs/2409.10580) - Guidelines for trustworthy medical AI systems
- <span class="year-entry">2024</span>[Mechanistic Interpretation through Contextual Decomposition in Transformers](https://arxiv.org/pdf/2407.00886) - Method for understanding transformer decision-making processes
- <span class="year-entry">2024</span>[LoRA+: Efficient Low Rank Adaptation of Large Models](https://arxiv.org/abs/2402.12354) - Improved technique for fine-tuning large language models
- <span class="year-entry">2024</span>[VDS book review](https://hdsr.mitpress.mit.edu/pub/5322mn1e/release/3) in Harvard Data Science Review - Praise for pedagogical approach to responsible data science
- <span class="year-entry">2024</span>[Berkeley-Stanford Workshop on Veridical Data Science](https://na.eventscloud.com/website/69057/) ([videos](https://www.youtube.com/playlist?list=PLKW2Azk23ZtSrg4jwLSioaV1bwmFnQ1Bx)) - Inaugural academic collaboration on trustworthy AI
- <span class="year-entry">2024</span>[Veridical Data Science book](https://mitpress.mit.edu/9780262049191/veridical-data-science/) released ([online](http://vdsbook.com/)) - Comprehensive guide to responsible data analysis and decision making
- <span class="year-entry">2023</span>[Explaining black box text modules in natural language with language models](https://arxiv.org/abs/2305.09863) - Natural language explanations for neural network text processing
- <span class="year-entry">2023</span>[Diagnosing transformers: illuminating feature space for clinical decision-making](https://arxiv.org/abs/2305.17588) - Interpretability methods for medical transformer applications
- <span class="year-entry">2023</span>[COPSS Distinguished Award and Lecture](https://ww2.amstat.org/meetings/jsm/2023/webcasts/index.cfm), JSM Toronto - Recognition for contributions to statistical science and AI

[View all papers →](https://scholar.google.com/citations?user=xT19Jc0AAAAJ&hl=en)